### 2017-04-06 paper/sexSelectionSpacingAppendix-ver1.tex
Further edits of theory part; moved tables out of doc; moved comparison back to appendix

>>>>

\subsection{Comparing the Hazard and Simple Models}

As detailed in the main paper, one of the main advantages of the hazard model 
approach is that it can be used to predict what completed fertility, number of abortions, 
and sex ratios will be once women are done with childbearing.
This cannot be done with the simple model because it does not predict fertility
progression, does not take into account censoring, and cannot capture if parents
change their use of sex selection within a spell.
We can, however, still compare differences in predicted sex ratios for 
\emph{individual} parities.%
\footnote{
The simple model uses only observed births and the predicted sex ratio is 
therefore simply what we observe for the relevant uncensored sample.
}

For each parity, the simple model is
\begin{equation}
Y_i = \gamma + \alpha' \mathbf{Z_i} + \beta' \mathbf{X}_{i} + \epsilon_i,
\end{equation}
where $Y$ is a dummy variable that takes the value 1 if the child born is a boy and zero 
if it is a girl, and the vectors of explanatory variables, $\mathbf{Z}_{i}$ and 
$\mathbf{X}_{i}$, are the same as for the proposed method, except that 
$\mathbf{Z}_{i}$ obviously does not include the baseline hazard and that there are 
no time dependence.%
\footnote{
To ease presentation, the indicator for parity number is not shown.
}
I estimate this using a logit model, although a linear probability model leads to
similar results.
This model is estimated on all observed births for a given parity in the relevant 
period to ensure that the results are as close to the standard method of 
examining sex selection \citep[see, for example, ][]{retherford03b,jha06,abrevaya09}.
\begin{footnotesize}
\caption{Comparison of Hazard and Simple Models for  \protect\linebreak
Women with 8 or More Years of Education}
\label{tab:compare2}
\begin{tabular} {@{} l D{.}{.}{5.2} D{.}{.}{5.2} D{.}{.}{5.2} D{.}{.}{5.2}   @{}} \toprule
                                  &  \multicolumn{2}{c}{Urban}        & \multicolumn{2}{c}{Rural}         \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5}
                                  & \mco{Simple}     &  \mco{Hazard}   &  \mco{Simple}    & \mco{Hazard}      \\   \midrule
                                  & \multicolumn{4}{c}{1985--1994 - Second Birth} \\ \cmidrule(l){2-5}
                                  & \multicolumn{4}{c}{1 girl} \\
Number of women\tnote{a}          &             & 4,869   &         & 3,105   \\
Births\tnote{b}                   &     3,848   & 3,926   & 2,417   & 2,795   \\
Percentage boys                   &      54.0   &  54.7   &  52.6   &  53.5   \\
Sex ratio (boys per 100 girls)    &     117.4   & 120.7   & 110.9   & 115.1   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1985--1994 - Third Birth}  \\ \cmidrule(l){2-5}
%                                   & \multicolumn{4}{c}{1 boy and 1 girl}        \\
% Number of women\tnote{a}          &             & 3,460   &         & 1,987   \\
% Births\tnote{c}                   &     1,431   & 1,558   & 1,107   & 1,336   \\
% Percentage boys                   &      53.0   &  52.5   &  53.5   &  54.6   \\
% Sex ratio (boys per 100 girls)    &     112.9   & 110.4   & 115.0   & 120.2   \\
% \addlinespace
                                  & \multicolumn{4}{c}{2 girls}        \\
Number of women\tnote{a}          &             & 1,654   &         &   982   \\
Births\tnote{c}                   &     1,034   & 1,092   &   682   &   853   \\
Percentage boys                   &      60.8   &  63.7   &  54.7   &  57.4   \\
Sex ratio (boys per 100 girls)    &     155.3   & 175.3   & 120.7   & 134.6   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1995--2006 - Second Birth} \\ \cmidrule(l){2-5}
                                  & \multicolumn{4}{c}{1 girl} \\                                                  
Number of women\tnote{a}          &             & 3,774   &         & 2,823   \\
Births\tnote{b}                   &     3,031   & 2,704   & 2,489   & 2,407   \\
Percentage boys                   &      58.0   &  58.6   &  56.0   &  56.9   \\
Sex ratio (boys per 100 girls)    &     138.1   & 141.3   & 127.1   & 132.2   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1995--2006 - Third Birth}  \\ \cmidrule(l){2-5}
%                                   & \multicolumn{4}{c}{1 boy and 1 girl}        \\
% Number of women\tnote{a}          &             & 2,357   &         & 1,819   \\
% Births\tnote{c}                   &       830   &   775   &   853   &   998   \\
% Percentage boys                   &      53.9   &  52.9   &  55.5   &  56.5   \\
% Sex ratio (boys per 100 girls)    &     116.7   & 112.1   & 124.5   & 129.8   \\
% \addlinespace
                                  & \multicolumn{4}{c}{2 girls}        \\
Number of women\tnote{a}          &             & 1,000   &         &   863   \\
Births\tnote{c}                   &       695   &   600   &   690   &   661   \\
Percentage boys                   &      65.5   &  62.4   &  62.0   &  61.7   \\
Sex ratio (boys per 100 girls)    &     189.6   & 166.0   & 163.4   & 161.0   \\
\item \hspace*{-0.5em} \textbf{Note.} 
The simple models are estimated using logit on all births for a given parity that
occurred during the relevant time period.
The hazard models are estimated using all spells that began in the relevant time period.
For the second birth the hazard model covers the period from beginning of spell to 6 
years (24 quarters) after the birth of the first child.
For the third birth it covers the period from beginning of spell to 7.25 years 
(29 quarters) after the birth of the second child for the period 1985--1994 and from 
beginning of spell to 5.75 years (23 quarters) after the birth of the second child for 
the period 1995--2006.
\item[a] Number of women in period who began the spell with given sex composition of prior 
child/children in the relevant period.
\item[b] For Logit model the number of women who are observed to have a second birth
in the period. 
For hazard model the predicted number of second births that will occur between beginning 
of the spell and 6 years (24 quarters) after the birth of the first child.
% \item[c] Number of women in period who had one boy and one girl as their first two children.
\item[c] For Logit model the number of women who are observed to have a third birth
in the period. 
For hazard model the predicted number of third births that will occur between beginning of 
the spell and 7.25 years (29 quarters) after the birth of the second child for the
1985--1994 period and between beginning of the spell and 5.75 years (23 quarters) after 
the birth of the second child for the 1995--2006 period.
\end{footnotesize}

I focus on those cases where there is evidence of sex-selective abortions: the spells
from parity one to two and two to three for women with eight or more years of education.%
\footnote{
The results for the spell from parity three to four are available on request.
They are not presented because the number of educated women who end up with three births 
and the probability of having a fourth birth are small, which makes
comparing the simple and hazard models with any precision difficult.
}
Table \ref{tab:compare2} shows the number of women who enter the second and third spells
for the hazard model,
the number of births, observed in the case of the simple model and predicted in the case 
of the hazard model, predicted percent boys, and the associated predicted sex ratio.
It is possible for the number of births used for the simple model to be larger than the
hazard models' predicted number of births because the simple model is based
on all births that occurred in the period---independently of when the prior birth
occurred---whereas the hazard model uses only spells that began in the period.

For the 1985--1994 period---which is when sex selection become available---the 
hazard model consistently predicts a higher sex ratio than the logit model.
Furthermore, some of these differences are substantial.
For urban women with 2 girls as their first two children, the hazard model
predicts a sex ratio of 175 boys per 100 girls, whereas the simple method
``only'' predicts 155 boys per 100 girls.
The difference is smaller in rural areas but still 14 boys per 100 girls
higher for the hazard model than for the simple model.
A similar, if less pronounced pattern, show up for the second birth,
where the hazard model predict 3 and 5 boys more per 100 girls for
urban and rural women.

For the latest period, 1995--2006, the hazard model still produces higher 
predicted sex ratios for second birth and the differences are of a 
similar magnitude to the 1985--1994 period.
The predicted sex ratios for the third birth are, however, 
higher for the simple model than for the hazard model.
This difference is substantially for urban women where the difference 
is 23 boys extra per 100 girls.

A possible explanation for these differences is that the simple model 
fails to capture the move towards lower fertility and the subsequent 
increase in the use of sex-selective abortions at earlier parities that 
occurred over the periods.
The results also suggest that the take-up of sex selection occurred 
rapidly as the new methods became available and that this take-up
is not fully captured by the simple model.
More broadly, a possible interpretation of the differences is that
the simple model will tend to underestimate when sex-selective abortions
are initially spreading in use and overestimate when the spread of sex 
selection is slowing.
This is most clearly seen for the third births for urban women where
the hazard model show a \emph{decline} is use of sex-selective abortions,
consistent with the lower fertility as discussed in the main paper,
while the simple model show a substantial increase.
